---
title: "DOHMH New York City Restaurant Inspection Results"
author: "LeTicia Cancel"
date: "today()"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Update with updated data using the API https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j 

Create visualization using Tableau https://support.socrata.com/hc/en-us/articles/115011744048 

# Introduction

This dataset is publicly available. It contains the restaurant inspection results from XX to 'max_date'. This dataset only includes active restaurants so the data will change over time. It also includes restaurants that have applied for a permit but have not yet been inspected. 

New libraries
```{r}
library(RSocrata)
library(dplyr)
library(tidyverse)
library(gridExtra)
library(conflicted)
```

## Get Dataset

```{r}
#get dataset using API
token <- "ctZLVtoG6N0anBu9lXENnsz8S"
df <- read.socrata("https://data.cityofnewyork.us/resource/43nn-pn8j.json", app_token = token)
```


```{r}
#get dataset using local csv file
#file is too large for github so it will not be uploaded to the respository
#df <- read.csv("DOHMH_New_York_city_Restaurant_Inspection_Results.csv")
```


```{r}
#get dataset from excel file on Github
```

## Data Exploration


Explore the data
```{r}
head(df)
nrow(df)
str(df)
summary(df)
```

Earliest inspection and most recent inspection. Active companies that have not been inspected yet all have an inspection date of 1900-01-01 EST so we will add a new column with a flag to distinguish between companies waiting for an inspection ("pending") and those who had an inspection ("complete"). 
```{r}
min(df$inspection_date)
max_date <- max(df$inspection_date)
```

Two new datasets are made with the new Status flag. 209,696 companies have a completed inspection and 2,544 companies are pending. 
```{r}
df_update <- df %>%
  mutate(status = if_else(inspection_date == "1900-01-01 EST", "pending", "complete"))

# companies with completed inspections
df_complete <- df_update %>%
  filter(status == "complete")

# companies pending inspection
df_pending <- df_update %>%
  filter(status == "pending")


nrow(df_complete)
nrow(df_pending)
```

The complete dataframe has NA's in the Grade Data columnn but not in the Grade column. We would assume that if a company with a Grade should also have a Grade Date so this is somethign we will have to explore further. 
```{r}
summary(df_complete)
```

Frequency tables

Boroughs
```{r}
#complete list
p_boro_complete <- ggplot(df_complete, aes(x = boro)) +
  geom_bar() + 
  coord_flip()

#pending list
p_boro_pend <- ggplot(df_pending, aes(x = boro)) +
  geom_bar() +
  coord_flip()

grid.arrange(p_boro_complete, p_boro_pend, ncol=2)

```

Cuisine Description - there are too many options so a bar plot is not useful. need another visualization
```{r}
#complete list
p_boro_complete <- ggplot(df_complete, aes(x = cuisine_description)) +
  geom_bar() + 
  coord_flip()

#pending list
p_boro_pend <- ggplot(df_pending, aes(x = boro)) +
  geom_bar() +
  coord_flip()

grid.arrange(p_boro_complete, p_boro_pend, ncol=2)

ggplot(df_complete, aes(x = violation_code)) +
  geom_bar() 
```

Violation code - there are too many options so a bar plot is not useful. need another visualization
```{r}
#complete list
p_boro_complete <- ggplot(df_complete, aes(x = cuisine_description)) +
  geom_bar() + 
  coord_flip()

#pending list
p_boro_pend <- ggplot(df_pending, aes(x = boro)) +
  geom_bar() +
  coord_flip()

grid.arrange(p_boro_complete, p_boro_pend, ncol=2)

ggplot(df_complete, aes(x = violation_code)) +
  geom_bar() 
```


Critical Flag - only one chart is needed since this column is only applicable to restaurants with an inspection. 
```{r}
ggplot(df_complete, aes(x = boro, fill = critical_flag)) +
  geom_bar() + coord_flip()
```

Scores - violin plot after converting the scores to a numeric data type 
```{r}
score <- df_complete %>%
  select(boro,score)

score$score <- as.numeric(score$score)

min(score$score)
max(score$score)

score %>%
  ggplot(aes(x = boro, y = score)) +
  geom_violin()
```

# Old Code



Libraries needed for analysis.
```{r message=FALSE, warning=FALSE}
library(sf)
library(DBI)
library(RCurl)
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Data 1 - NYC Restaurant Inspection 

### Get Dataset 

The first data set that will be analyzed are all of the NYC Restaurant Inspection results from 2019 to 10/1/2020. The data will be pulled from a csv file that is on GitHub.
```{r message=FALSE,warning=FALSE}
#get URL and import data from GitHub
site <- "https://raw.githubusercontent.com/ltcancel/Project2_Data607_F20/master/DOHMH_New_York_City_Restaurant_Inspection_Results.csv"
df <- read_csv(url(site))

head(df)
```
### Tidy the data/Analysis

Q. How many violations were issued for restaurants in each borough for each year? 
```{r}
#separate date into its own columns
df <- df %>%
  separate('INSPECTION DATE',sep="/",into = c("month","day","year"),remove = FALSE)

#not every row has a value for BORO so I want to exclude this data. I also want to include anyone who does not have a violation description
df1Filter <- df %>% group_by(year) %>% count(BORO) %>% filter(BORO != 0 & !is.na('VIOLATION DESCRIPTION'))

head(df1Filter)
```

After plotting the data, we can see that the number of violations has increased over the years. Manhattan consistently has the greatest number of violations. Brooklyn and Queens are very close, and Staten Island has the least number of violations. I think this data needs further analysis to see the percentage of restaurants with violations and the percentage of restaurants with an A grade. Manhattan may have a larger number of restaurants making their violations seem inflated when compared to the other boroughs. 

```{r}
ggplot(df1Filter) +
  geom_bar(mapping = aes(x=BORO, y = n, fill = BORO), stat = 'identity') +
  facet_wrap(~year) +
  coord_flip() +
  labs(title = "Resturant Insepction Violations",x="Borough",y="Violations")
```



